---
title: " Association"
author: "YI XIAO"
date: "3/13/2019"
output: pdf_document 
---
## AIM 1 
### a)	Scatter plot of hsTNI (or log hsTNI) vs. ACHDIARES
```{r include = FALSE}
knitr::opts_chunk$set(echo=FALSE)
```

```{r}
library(readxl)
##baseline 
data=read_xlsx('/Users/xiaoyi/Documents/2018Fall/Clinical/cardiac_MRI_and_troponin_and_BNP_levels/ahmedwisecvd3182019.xlsx')
 
aim1=data.frame(hsTNI=data$Mean_Conc__Pg_ml,ACHDIARES=data$ACHDIARES,ACHCBF_BEST=data$ACHCBF_BEST,LVEDP=data$LVEDP)


plot(aim1$ACHDIARE,log(aim1$hsTNI),main=' log hsTNI vs. ACHDIARES ',xlab='ACHDIARE',ylab='log hsTNI')
abline(v=0,col='red',lty=3)
plot(aim1$ACHDIARE,aim1$hsTNI,main='hsTNI vs. ACHDIARES ',xlab='ACHDIARE',ylab='hsTNI')
abline(v=0,col='red',lty=3)



```

### b)	Scatter plot of hsTNI (or log hsTNI) vs. ACHCBF_BEST 
```{r}
plot(aim1$ACHCBF_BEST,log(aim1$hsTNI),main=' log hsTNI vs.ACHCBF_BEST ',xlab='ACHCBF_BEST ',ylab='log hsTNI')
abline(v=20,col='red',lty=3)
abline(v=50,col='red',lty=3)
plot(aim1$ACHCBF_BEST,aim1$hsTNI,main='hsTNI vs. ACHCBF_BEST ',xlab='ACHCBF_BEST',ylab='hsTNI')
abline(v=20,col='red',lty=3)
abline(v=50,col='red',lty=3)
```






### c)	Compare hsTNI (or log hsTNI) between those with ACHDIARES<0 vs. ACHDIARES>=0
```{r}
aim1$bi_ACHDIARES=factor(ifelse(aim1$ACHDIARES>0,1,0))
 wilcox.test( hsTNI~bi_ACHDIARES , data = aim1,
                   exact = FALSE)

```

### d)	Compare hsTNI (or log hsTNI) between those with ACHCBF_BEST < 20 vs. ACHCBF_BEST >=20 using Wilcoxon test
```{r}
aim1$bi_ACHCBF_BEST20=factor(ifelse(aim1$ACHCBF_BEST>20,1,0))
 wilcox.test( hsTNI~bi_ACHCBF_BEST20 , data = aim1,
                   exact = FALSE)

```

### e)	Compare hsTNI (or log hsTNI) between those with ACHCBF_BEST < 50 vs. ACHCBF_BEST >=50 using Wilcoxon test
```{r}
aim1$bi_ACHCBF_BEST50=factor(ifelse(aim1$ACHCBF_BEST>50,1,0))
 wilcox.test( hsTNI~bi_ACHCBF_BEST50 , data = aim1,
                   exact = FALSE)

```


### f)	Compare hsTNI (or log hsTNI) between those with ACHCBF_BEST < median vs. ACHCBF_BEST >=median using Wilcoxon test

```{r}
me=median(aim1$ACHCBF_BEST,na.rm=TRUE)
aim1$bi_ACHCBF_BEST_me=factor(ifelse(aim1$ACHCBF_BEST>me,1,0))
 wilcox.test( hsTNI~bi_ACHCBF_BEST_me , data = aim1,
                   exact = FALSE)

```
### boxplot on  ACHCBF_BEST
```{r}
aim1$lg_hsTNI=log(aim1$hsTNI)
library("ggpubr")
ggboxplot(aim1, x = "bi_ACHCBF_BEST50", y = "lg_hsTNI", 
          color = "bi_ACHCBF_BEST50", palette = c("#00AFBB", "#E7B800"),
          ylab = "log_hsTNI", xlab = "ACHCBF_BEST > 50")
```
```{r}
ggboxplot(aim1, x = "bi_ACHCBF_BEST20", y = "lg_hsTNI", 
          color = "bi_ACHCBF_BEST20", palette = c("#00AFBB", "#E7B800"),
          ylab = "log_hsTNI", xlab = "ACHCBF_BEST > 20")
```
```{r}
ggboxplot(aim1, x = "bi_ACHCBF_BEST_me", y = "lg_hsTNI", 
          color = "bi_ACHCBF_BEST_me", palette = c("#00AFBB", "#E7B800"),
          ylab = "log_hsTNI", xlab = "ACHCBF_BEST > median")
```

### h) Scatter plot of hsTNI (or log hsTNI) vs. LVEDP

```{r}
plot(aim1$LVEDP,aim1$hsTNI,main='hsTNI vs. LVEDP ',xlab='LVEDP',ylab='hsTNI')
abline(v=25,col='red',lty=3)
plot(aim1$LVEDP,log(aim1$hsTNI),main='log hsTNI vs. LVEDP ',xlab='LVEDP',ylab='log hsTNI')
abline(v=25,col='red',lty=3)
```

### i) Compare hsTNI (or log hsTNI) between those with  LVEDP < 25 vs.  >=25 using Wilcoxon  test

```{r}
aim1$bi_LVEDP=factor(ifelse(aim1$LVEDP>25,1,0))
 wilcox.test( hsTNI~ bi_LVEDP, data = aim1,exact = FALSE)

```




## AIM 2 
### a)	Scatter plot of hsTNI (or log hsTNI) vs. each cMRI variable 
```{r}
col=c('longstrain','LSRs','LSRd1','LSRd2','radstrain','RSRs','RSRd1',	'RSRd2','circstrainpercentage','CSRs','CSRd1','CSRd2','Mean_WMPRI')


 
aim2=data[,col]
aim2$lg_hsTNI=log(data$Mean_Conc__Pg_ml)

#plot(log(aim2$hsTNI)~.,data=aim2,ylab='log hsTNI')
```
```{r}
library(tidyr)
aim2 %>%
  gather(-lg_hsTNI, key = "var", value = "value") %>%
  ggplot(aes(y =  lg_hsTNI, x = value)) +
    geom_point(cex=0.5) +
    facet_wrap(~ var, scales = "free") +
    theme_bw()
```



###  b)	Correlation matrix of all cMRI variables  

```{r}
cMRI=aim2[,-ncol(aim2)]
res=cor(cMRI,method = "spearman",use="complete.obs")
```
```{r}
library(corrplot)
corrplot(res, type = "upper", order = "original", 
         tl.col = "black", tl.srt = 45)
```

### change order : group them by cardiac cycle phase:


```{r}
col=c('LSRs','RSRs','CSRs','LSRd1','RSRd1','CSRd1','LSRd2','RSRd2','CSRd2','longstrain','radstrain','circstrainpercentage')
aim2=data[,col]
aim2$hsTNI=data$Mean_Conc__Pg_ml

```
```{r}
cMRI=aim2[,-ncol(aim2)]
res=cor(cMRI,method = "spearman",use="complete.obs")
```
```{r}
library(corrplot)
corrplot(res, type = "upper", order = "original", 
         tl.col = "black", tl.srt = 45)
```



### c)	Determine PCA or FA on cMRI variables
```{r}
library(visdat)
vis_miss(cMRI)
```
```{r}
na.omit(cMRI)->cMRI.omit
bio.pca=prcomp(cMRI.omit,
                 center = TRUE,
                 scale. = TRUE)
plot(bio.pca,type = "l")
```
```{r}
VE <- bio.pca$sdev^2
PVE <- VE / sum(VE)
round(PVE, 2)
```
```{r}
bio.pca$rotation[,1:4]
```


### d)	Extract principal components of factors and then examine the association with hsTNI and BNP (possibly adjusted for age)

```{r}
library(nFactors)
my.data=cMRI.omit
ev <- eigen(cor(my.data)) # get eigenvalues
ap <- parallel(subject=nrow(my.data),var=ncol(my.data), rep=100, cent=.05)
nS <- nScree(x=ev$values, aparallel=ap$eigen$qevpea)
plotnScree(nS)
```
 The Kaiser-Guttman rule says that we should choose all factors with eigenvalue greater than 1, so we choose n=4
 
```{r}
ev$values
```
 
```{r}
fit <- factanal(my.data, 
                4,              # number of factors to extract
                rotation="varimax")
load<-fit$loadings
load
```
 

 
#### adjust for log(BNP)
```{r}
col=c('radstrain','RSRs','RSRd1',	'RSRd2','circstrainpercentage','CSRs','CSRd1','CSRd2',	'longstrain','LSRs','LSRd1','LSRd2','Mean_WMPRI')
aim2=data[,col]
aim2$hsTNI=data$Mean_Conc__Pg_ml
aim2$bnp=data$bnp
aim2$age=data$age
# adjust for  Diabetes
aim2$HSDIB=data$HSDIB
# adjust for Hypertension 
aim2$HSHYP=data$HSHYP
# adjust for Dyslipidemia
aim2$DISLP=data$DISLP
aim2.omit=na.omit(aim2)
fit <- principal(aim2.omit[,col], nfactors=4, rotate="varimax")
fit1=lm(fit$scores[,1]~log(bnp)+age+HSHYP+HSDIB+DISLP,data=aim2.omit)
fit2=lm(fit$scores[,2]~log(bnp)+age+HSHYP+HSDIB+DISLP,data=aim2.omit)
fit3=lm(fit$scores[,3]~log(bnp)+age+HSHYP+HSDIB+DISLP,data=aim2.omit)
fit4=lm(fit$scores[,4]~log(bnp)+age+HSHYP+HSDIB+DISLP,data=aim2.omit)
fit.BNP=list(fit1,fit2,fit3,fit4)
```
```{r}
library(sjPlot)

tab_model(fit.BNP,dv.labels=c('factor1','factor2','factor3','factor4'),show.se = TRUE)
```
#### adjust for log(hsTNI)
```{r}
fit1=lm(fit$scores[,1]~log(hsTNI)+age+HSHYP+HSDIB+DISLP,data=aim2.omit)
fit2=lm(fit$scores[,2]~log(hsTNI)+age+HSHYP+HSDIB+DISLP,data=aim2.omit)
fit3=lm(fit$scores[,3]~log(hsTNI)+age+HSHYP+HSDIB+DISLP,data=aim2.omit)
fit4=lm(fit$scores[,4]~log(hsTNI)+age+HSHYP+HSDIB+DISLP,data=aim2.omit)
fit.hsTNI=list(fit1,fit2,fit3,fit4)
```

```{r}
library(sjPlot)

tab_model(fit.hsTNI,dv.labels=c('factor1','factor2','factor3','factor4'),show.se = TRUE)
```
## AIM 3 
### explore if women with higher urine hsTNI at baseline have worsened LV strain rate on follow up cardiac MRI 
```{r}
data_foll=read_xlsx('/Users/xiaoyi/Documents/2018Fall/Clinical/cardiac_MRI_and_troponin_and_BNP_levels/ahmedwisecvd3182019.xlsx',sheet = 'Follow Up CMRI')

col=c('LSRs','RSRs','CSRs','LSRd1','RSRd1','CSRd1','LSRd2','RSRd2','CSRd2','longstrain','radstrain','circstrainpercentage','cmerg')
aim3=data_foll[,col]
aim2$cmerg=data$cmerg
aim.all<-merge(aim2,aim3,by='cmerg')
# the sequence of data should be aim3-aim2 (follow -baseline )
M <- merge(aim3,aim2,by='cmerg')

S <- M[,grepl("*\\.x$",names(M))] - M[,grepl("*\\.y$",names(M))]

minus.aim<-cbind(M[,1,drop=FALSE],S)

df=data.frame(lg_hsTNI=log(aim2$hsTNI),cmerg=aim2$cmerg)
aim.all<-merge(minus.aim,df,by='cmerg')
cov=c('age','HSHYP','HSDIB','DISLP','cmerg')
#use baseline covariate !!
aim.all.cov<-merge(data[,cov],aim.all,by='cmerg')
aim.all.new<-aim.all[,-which(colnames(aim.all)=='cmerg')]
colnames(aim.all.new)=gsub(".x$",'',colnames(aim.all.new))
colnames(aim.all.cov)=gsub(".x$",'',colnames(aim.all.cov))

```


### plot difference of follow-up and baseline vs. baseline hsTIN
```{r}
library(tidyr)
aim.all.new %>%
  gather(-lg_hsTNI, key = "var", value = "value") %>%
  ggplot(aes(y = value, x = lg_hsTNI)) +
    geom_point(cex=0.5) +
    facet_wrap(~ var, scales = "free") +
    theme_bw()
```

### regression model
```{r}
col=c('radstrain','RSRs','RSRd1',	'RSRd2','circstrainpercentage','CSRs','CSRd1','CSRd2',	'longstrain','LSRs','LSRd1','LSRd2')
fit.all=list()


for (i in 1:length(col)){
  fit=lm(get(col[i])~lg_hsTNI+age+HSHYP+HSDIB+DISLP,data=aim.all.cov)
  fit.all[[i]]=fit
}

```

```{r}
library(sjPlot)

tab_model(fit.all,dv.labels=col,show.se = TRUE)
```

